---
title: "First wave vs. second wave"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

library(tidyverse)
library(ggplot2)
library(scales)
library(dplyr)

```

# Visualise the number of confirmed cases database

## Extract data from file "monthly_covid_data.xlsx" to establish a database for the number of confirmed cases in selected countries.
```{r}
rm(list=ls())

confirmed_case <- readxl::read_xlsx("data/monthly_covid_confirmed_cases_per1000_selected.xlsx")

confirmed_case <- as.data.frame(confirmed_case[1:10,1:12])

colnames(confirmed_case)[1] <- c("month")
confirmed_case[,1] <- c("01","02","03","04","05","06","07","08","09","10")

# Convert the dataframe into tidy form
tidy_confirmed_case  <- confirmed_case %>% 
  pivot_longer(cols="South_Africa":"Australia",
               names_to = "country", 
               values_to = "confirmed_cases_per_100k")

```

## Identify countries that are experiencing the second wave
```{r}
# plot the confirmed cases in several countries around the world
ggplot(tidy_confirmed_case, 
       aes(x=month, 
           y=confirmed_cases_per_100k, 
           group = country, 
           color = country)) + 
  geom_line(size=1)+
  theme_bw()+
  labs(
    title = "Monthly confirmed cases per country",
    y="confirmed cases per 100k",
    color= "Countries"
  )+
  xlab("")+
  scale_y_continuous(label = comma) + 
  scale_x_discrete(labels = c("Jan","Feb","Mar","April","May","Jun","Aug","Sep","Oct","Nov")) +
  theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    plot.margin = unit(c(0, 0, 0, 0), "cm"), 
        strip.text.x = element_text(size = 11, face = "bold", vjust =1), 
        title =element_text(size=16, face='bold', vjust =1, ), 
        plot.subtitle =element_text(size=14, vjust =1), 
        axis.text.x = element_text(color="black", size=10, face="plain"),
        axis.text.y= element_text(color="black", size=11, face="plain"), 
        panel.spacing = unit(2, "lines")
  )

# exclude countries that are still experiencing the first wave, plot again
confirmed_case_selected <-
  confirmed_case[,c("month","Italy","France","Spain","Germany","UK","Belgium")]

tidy_confirmed_case_selected  <- confirmed_case_selected %>% 
  pivot_longer(cols="Italy":"Belgium",
               names_to = "country", 
               values_to = "confirmed_cases_per_100k")

ggplot(tidy_confirmed_case_selected, 
       aes(x=month, y=confirmed_cases_per_100k, group = country, color = country)) + 
  geom_line(size=1)+
  theme_bw()+
  labs(
    title = 
    "Monthly confirmed cases in countris under 
    the second wave",
    y="confirmed cases per 100k",
    color= "Countries"
  )+
  xlab("")+
  scale_y_continuous(label = comma) +
  scale_x_discrete(labels = c(
    "Jan","Feb","Mar","April","May","Jun","Aug","Sep","Oct","Nov")) +
  theme_minimal()+
  theme(
    axis.title.x=element_blank(),
    legend.title = element_blank(),
    plot.margin = unit(c(0, 0, 0, 0), "cm"), 
        strip.text.x = element_text(size = 11, face = "bold", vjust =1), 
        title =element_text(size=16, face='bold', vjust =1, ), 
        plot.subtitle =element_text(size=14, vjust =1), 
        axis.text.x = element_text(color="black", size=10, face="plain"),
        axis.text.y= element_text(color="black", size=11, face="plain"), 
        panel.spacing = unit(2, "lines"),
    legend.position = "bottom"
  )
```



# create the databse of the absolute number of airflights per route
```{r}
airpo_info <- as.data.frame(
  readxl::read_xlsx(
    "data/Num_flights_per_route_per_month_top10kroutes_airport_info.xlsx")[,-2])

colnames(airpo_info)[1] <- 'Route'

routes_info <- as.data.frame(
  readxl::read_xlsx(
    "data/Num_flights_per_route_per_month_top10kroutes_wide.xlsx")[13:22,])

routes_info[,1] <- c("01","02","03","04","05","06","07","08","09","10")

tidy_routes_info  <- routes_info %>% 
  pivot_longer(
               cols=2:10001,
               names_to = "Route", 
               values_to= "number_of_flights"
  ) 
colnames(tidy_routes_info)[1] <- 'Month'


combined_airport_info <- left_join(tidy_routes_info,
                           airpo_info,
                           by=c("Route"))

combined_airport_info$number_of_flights <- as.numeric(
  combined_airport_info$number_of_flights)
combined_airport_info$orig_ctry <- as.character(combined_airport_info$orig_ctry)

```

## total number of routes per country
```{r}
flights_number_per_country <- combined_airport_info %>% 
  group_by(Month, orig_ctry) %>% 
  summarise(monthly_volume = sum(number_of_flights))

flights_number_selected <- flights_number_per_country %>% 
  filter(orig_ctry=="FR"|
           orig_ctry=="GB"|
           orig_ctry=="ES"|
           orig_ctry=="DE"|
           orig_ctry=="US"|
           orig_ctry=="IN") 
colnames(flights_number_selected)[1] <- 'month'
colnames(flights_number_selected)[2] <- 'country'

# rebase the number by setting the data in 2020 Jan as 100
rebase_benchmark <- flights_number_selected %>% 
  group_by(country) %>% 
  filter (month == "01") %>% 
  select (country, monthly_volume) %>% 
  mutate (rebasenumber = monthly_volume) %>% 
  select (country, rebasenumber)

rebase_benchmark1 <- left_join(flights_number_selected,
                           rebase_benchmark,
                           by=c("country"))

rebased_flight_number <- rebase_benchmark1 %>% 
  mutate (rebased_number_of_flight = monthly_volume / rebasenumber ) %>% 
  select (month, country, rebased_number_of_flight)
```

# generate combined dataset for next step visualisation
```{r}
countries_selected <-
  confirmed_case[,c("month","India","France","Spain","Germany","UK","US")]

countries_selected_tidy <- countries_selected %>% 
  pivot_longer(
               cols="India":"US",
               names_to = "country", 
               values_to= "confirmed_cases_per_100k"
  ) 

rebased_flight_number$country[rebased_flight_number$country=="DE"] <- "Germany"
rebased_flight_number$country[rebased_flight_number$country=="FR"] <- "France"
rebased_flight_number$country[rebased_flight_number$country=="ES"] <- "Spain"
rebased_flight_number$country[rebased_flight_number$country=="GB"] <- "UK"
rebased_flight_number$country[rebased_flight_number$country=="IN"] <- "India"

selected_countries_covid <-
  confirmed_case[,c("month","US","France","Spain","Germany","India","UK")]

selected_countries_covid_tidy  <- selected_countries_covid %>% 
  pivot_longer(cols="US":"UK",
               names_to = "country", 
               values_to = "confirmed_cases_per_100k")

combined_data <- left_join(rebased_flight_number,
                           selected_countries_covid_tidy,
                           by=c("month","country"))

combined_data$rebased_number_of_flight <- as.numeric(
  combined_data$rebased_number_of_flight)
combined_data$confirmed_cases_per_100k <- as.numeric(
  combined_data$confirmed_cases_per_100k)

# separate data set to see countries that are experiencing different stages of COVID
combined_data_2nd_wave <- data.frame(combined_data %>% 
  filter((country=="France"|
           country=="UK"|
           country=="Spain"|
           country=="Germany")))

combined_data_india <- data.frame(combined_data %>% 
  filter((country=="India")))

combined_data_us <- data.frame(combined_data %>% 
  filter((country=="US")))

```


# final visualise - contrast the trends

## plot for countries that are experiencing the second wave
```{r fig.width=12, fig.height=8}
# rescale this variable
#plot
ggplot(data = combined_data)+
  geom_line(aes(x = month, 
                y = rebased_number_of_flight,
                group = country,
                color = "blue"),
            # color = "#001F62",
            size=0.8) +
  geom_line(aes(x = month, 
                y = confirmed_cases_per_100k/2800, # rescale
                group = country,
                color = "red"),
            # color = "#c6102f",
            size=0.8) +
  facet_wrap(~country)+
  scale_y_continuous(limits = c(0,1.3),
                     label = percent,
                           sec.axis = sec_axis(~.*2800, #divide the number multiplied in rescale step
                                               name = "number of cases per 100k",))+
  scale_x_discrete(labels = c("Jan","Feb","Mar","April","May","Jun","Aug","Sep","Oct","Nov"))+
  geom_rect(aes(xmin=3,xmax=4,ymin=-Inf, ymax=1.3),
            fill='#E5E5E5',
            alpha=0.03)+
  geom_rect(aes(xmin=9,xmax=10,ymin=-Inf, ymax=1.3),
            fill='#E5E5E5',
            alpha=0.03)+
  ggtitle("Contrast the trend of confirmed cases and the number of flights",
          "How they are changing from the first peak to now?")+
  scale_color_manual(values= c("#001F62","#c6102f"), labels=c("Number of flights per month","Daily confirmed cases per 100k"))+
  geom_hline(yintercept=1, 
             color="#B6B6B6", 
             size = 0.5,
             linetype = 5) +
  theme_minimal()+
  theme(
        axis.title.x = element_blank(),
        legend.text = element_text(color="black", size=12, face="plain"), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"), 
        strip.text.x = element_text(size = 11, face = "bold", vjust =1), 
        title = element_text(size=16, face='bold', vjust =1, ), 
        plot.subtitle = element_text(size=14, vjust =1), 
        axis.text.x = element_text(color="black", size=11, face="plain"),
        axis.text.y= element_text(color="black", size=11, face="plain"), 
        panel.spacing = unit(2, "lines"),
        legend.title = element_blank(),
        legend.position = "bottom")+
  ylab("change in traffic volume")

```
